---
title: "raytracing"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[![Travis-CI Build Status](https://img.shields.io/travis/com/salvatirehbein/raytracing?style=for-the-badge)](https://travis-ci.com/github/salvatirehbein/raytracing)
[![Codecov test coverage](https://codecov.io/gh/salvatirehbein/raytracing/branch/master/graph/badge.svg)](https://codecov.io/gh/salvatirehbein/raytracing?branch=master)

## Raytracing Documentation

```raytracing``` package uses Hoskins and Ambrizzi (1993) equations and a similar methodology of Yang and Hoskins (1996). This document describes how does the ```raytracing``` R package works with examples. 

It is required an initial coordinate pairs that represents the source for the Rossby wave.

## Including Code

The ```raytracing``` include 4 main functions inside the principal program, ```ray```:

```{r funcoes, echo = FALSE}
library(knitr)
df <- data.frame(
  Functions = c("calcUg", "calcVg", "betaks", "ray", "ypos"),
  Arguments = c("betamz, umz, y, lat, K, w, a", 
                "betamz, umz, y, lat, K, direction, tl, a",
                "ifile, varname, latname, lonname, ofile, a, plots, show.warnings",
                "betamz, umz, lat, x0, y0, K, dt,  itime, direction, tl, a",
                "y, lat"), 
  Description = c("Calculates the zonal group velocity", 
                  "Calculates the meridional group velocity",
                  "Calculates the stationary Rossby wave number, meridional gradient of the absolute vorticity, and zonal wind in mercator coordinates",
                  "Calculates the Rossby wave ray paths",
                  "Obtain the closest position of a latitude in a vector")
)
knitr::kable(x = df, caption = "**Table** **1** **-** ```raytracing``` functions and arguments.")
```

## Example

The Rossby wave ray paths obtained in Coelho et al. (2015) case with the stationary Rossby wave number in background.

```{r onda, echo=FALSE, fig.height=4, fig.width=10, warning=FALSE, message=FALSE}
library(raytracing, quietly = TRUE)
input <- system.file("extdata",
                     "uwnd.mon.mean_200hPa_2014JFM.nc",
                     package = "raytracing")
b <- betaks(ifile = input)
a <- ray(betamz = colMeans(b$betam, na.rm = TRUE),
         umz = colMeans(b$um, na.rm = TRUE),
         lat = b$lat,
         K = 3,
         itime = 30,
         x0 = -135 + 360,
         y0 = -30,
         dt = 6 * 60 * 60,
         direction = -1,
         verbose = FALSE)
a$x2 <- a$x0 - 360
library(ggplot2)
ww <- map_data('world', ylim=c(-60,60))

ggplot() +  theme_bw() +
  scale_y_continuous(limits = c(-60,60)) +
  geom_polygon(data = ww,
               aes(x = long, y = lat, group = group),
               alpha = 0.0, col = "grey") +
  geom_point(data = a[!(a$tun_y0 == -1 |
                          a$tun_y1 == -1 | a$id == 0), ],
             aes(x = x2, y = y0), size = 3) +
  geom_line(data = a[!(a$tun_y0 == -1 |
                         a$tun_y1 == -1 | a$id == 0), ],
            aes(x = x2, y = y0)) +
  ggtitle("Coelho et al. (2015):") +
  ylab(NULL) + xlab(NULL)
```

